---
title: "Effect Generalization"
author: Apoorva Lal
date: "`r Sys.Date()`"
---

<!--- For HTML renders - selection from math_shortcuts.tex --->
`r if (!knitr:::is_latex_output()) '
$\\DeclareMathOperator*{\\argmin}{argmin}$
$\\newcommand{\\abs}[1]{\\left\\vert {#1} \\right\\vert}$
$\\newcommand{\\indep}{\\perp\\!\\!\\!\\perp}$
$\\newcommand{\\var}{\\mathrm{Var}}$
$\\newcommand{\\ephi}{\\varephilon}$
$\\newcommand{\\phii}{\\varphi}$
$\\newcommand{\\tra}{^{\\top}}$
$\\newcommand{\\sumin}{\\sum_{i=1}^n}$
$\\newcommand{\\sumiN}{\\sum_{i=1}^n}$
$\\newcommand{\\norm}[1]{\\left\\Vert{#1} \\right\\Vert}$
$\\newcommand\\Bigpar[1]{\\left( #1 \\right )}$
$\\newcommand\\Bigbr[1]{\\left[ #1 \\right ]}$
$\\newcommand\\Bigcr[1]{\\left\\{ #1 \\right \\}}$
$\\newcommand\\SetB[1]{\\left\\{ #1 \\right\\}}$
$\\newcommand\\Sett[1]{\\mathcal{#1}}$
$\\newcommand{\\Data}{\\mathcal{D}}$
$\\newcommand{\\Ubr}[2]{\\underbrace{#1}_{\\text{#2}}}$
$\\newcommand{\\Obr}[2]{ \\overbrace{#1}^{\\text{#2}}}$
$\\newcommand{\\sumiN}{\\sum_{i=1}^N}$
$\\newcommand{\\dydx}[2]{\\frac{\\partial #1}{\\partial #2}}$
$\\newcommand\\Indic[1]{\\mathbb{1}_{#1}}$
$\\newcommand{\\Realm}[1]{\\mathbb{R}^{#1}}$
$\\newcommand{\\Exp}[1]{\\mathbb{E}\\left[#1\\right]}$
$\\newcommand{\\Expt}[2]{\\mathbb{E}_{#1}\\left[#2\\right]}$
$\\newcommand{\\Var}[1]{\\mathbb{V}\\left[#1\\right]}$
$\\newcommand{\\Covar}[1]{\\text{Cov}\\left[#1\\right]}$
$\\newcommand{\\Prob}[1]{\\mathbf{Pr}\\left(#1\\right)}$
$\\newcommand{\\Supp}[1]{\\text{Supp}\\left[#1\\right]}$
$\\newcommand{\\doyx}{\\Prob{Y \\, |\\,\\mathsf{do} (X = x)}}$
$\\newcommand{\\doo}[1]{\\Prob{Y |\\,\\mathsf{do} (#1) }}$
$\\newcommand{\\R}{\\mathbb{R}}$
$\\newcommand{\\Z}{\\mathbb{Z}}$
$\\newcommand{\\wh}[1]{\\widehat{#1}}$
$\\newcommand{\\wt}[1]{\\widetilde{#1}}$
$\\newcommand{\\wb}[1]{\\overline{#1}}$
$\\newcommand\\Ol[1]{\\overline{#1}}$
$\\newcommand\\Ul[1]{\\underline{#1}}$
$\\newcommand\\Str[1]{#1^{*}}$
$\\newcommand{\\F}{\\mathsf{F}}$
$\\newcommand{\\ff}{\\mathsf{f}}$
$\\newcommand{\\Cdf}[1]{\\mathbb{F}\\left(#1\\right)}$
$\\newcommand{\\Cdff}[2]{\\mathbb{F}_{#1}\\left(#2\\right)}$
$\\newcommand{\\Pdf}[1]{\\mathsf{f}\\left(#1\\right)}$
$\\newcommand{\\Pdff}[2]{\\mathsf{f}_{#1}\\left(#2\\right)}$
$\\newcommand{\\dd}{\\mathsf{d}}$
$\\newcommand\\Normal[1]{\\mathcal{N} \\left( #1 \\right )}$
$\\newcommand\\Unif[1]{\\mathsf{U} \\left[ #1 \\right ]}$
$\\newcommand\\Bern[1]{\\mathsf{Bernoulli} \\left( #1 \\right )}$
$\\newcommand\\Binom[1]{\\mathsf{Bin} \\left( #1 \\right )}$
$\\newcommand\\Pois[1]{\\mathsf{Poi} \\left( #1 \\right )}$
$\\newcommand\\BetaD[1]{\\mathsf{Beta} \\left( #1 \\right )}$
$\\newcommand\\Diri[1]{\\mathsf{Dir} \\left( #1 \\right )}$
$\\newcommand\\Gdist[1]{\\mathsf{Gamma} \\left( #1 \\right )}$
$\\def\\mbf#1{\\mathbf{#1}}$
$\\def\\mrm#1{\\mathrm{#1}}$
$\\def\\mbi#1{\\boldsymbol{#1}}$
$\\def\\ve#1{\\mbi{#1}}$
$\\def\\vee#1{\\mathbf{#1}}$
$\\newcommand{\\Mat}[1]{\\mathbf{#1}}$
$\\newcommand{\\eucN}[1]{\\norm{#1}}$
$\\newcommand{\\lzero}[1]{\\norm{#1}_0}$
$\\newcommand{\\lone}[1]{\\norm{#1}_1}$
$\\newcommand{\\ltwo}[1]{\\norm{#1}_2}$
$\\newcommand{\\pnorm}[1]{\\norm{#1}_p}$
'`


We simulate some experimental data with a true treatment
effect, introduce selection bias using the `selF` selection function,
and attempt to recover the true treatment effect using the observed data.

```{r}
suppressPackageStartupMessages(library(causalTransportR))
set.seed(42)

# simulate RCT
treatprob <- 0.5
# workhorse
dgp <- function(n = 10000, p = 10, treat.prob = treatprob,
                # bounds of X
                Xbounds = c(-1, 1),
                # nonlinear heterogeneity
                tauF = function(x) 1 / exp(-x[3]),
                # nonlinear y0
                y0F = function(x) pmax(x[1] + x[2], 0) + sin(x[5]) * pmax(x[7], 0.5),
                # nonlinear selection
                selF = function(x) x[1] - 3 * x[3] + pmax(x[4], 0)) {
        X <<- matrix(runif(n * p, Xbounds[1], Xbounds[2]), n, p)
        a <<- rbinom(n, 1, treat.prob)
        # generate outcomes using supplied functions
        TAU <<- apply(X, 1, tauF)
        Y0 <- apply(X, 1, y0F)
        selscore <- apply(X, 1, selF)
        # outcome
        y <<- (a * TAU + Y0 + rnorm(n))
        # selection
        s <<- rbinom(n, 1, plogis(selscore)) |> as.logical()
        # set outcomes for s = 0 as missing
        y[s == 0] <<- NA
        cat("True effect\n")
        cat(mean(TAU))
}
# this function populates the global namespace using <<- ; not recommended for real use
dgp()

```

We've introduced systematic selection bias in the missigness of `y`. This is plausible in settings where the experimental sample is systematically different from the overall population, or if the missing 'cohort' was assigned later and we want to call the experiment earlier based on complete data.

### naive

Using the AIPW estimator naively on complete data is ignoring the missingness problem, and gives us the wrong answer (we know the true effect above is `r mean(TAU)`).

```{r}
cat("\n Naive \n")
naive_fit <- ateGT(y[s == 1], a[s == 1], X[s == 1, ],
        nuisMod = "rlm", target = "insample", noi = F
)
naive_fit %>% summary()
```

## corrected estimators

Next we demonstrate the use of each of the estimators implemented in
`ateGT` on the simulated data above.

### AISW

```{r}
cat("\nAISW \n")
aipwfit <- ateGT(y, a, X, s,
        treatProb = treatprob, nuisMod = "rlm",
        estimator = "AISW", target = "generalize", noi = F
)
aipwfit %>% summary
```

As with any reweighting estimator, it is a good idea to study balance. The
plot method for the `ateGT` function produces the following plot, which suggests
substantial improvements in balance thanks to selection weights.

```{r}
aipwfit %>% plot(X)
```


```{r}

cat("\nISW \n")
ipwfit <- ateGT(y, a, X, s,
        treatProb = treatprob, nuisMod = "rlm",
        estimator = "ISW", target = "generalize", noi = F
)
ipwfit %>% summary
```

```{r}

cat("\nOM \n")
omfit <- ateGT(y, a, X, s,
        treatProb = treatprob, nuisMod = "rlm",
        estimator = "OM", target = "generalize", noi = F
)
omfit %>% summary

```

```{r}

cat("\nACW \n")
acwfit <- ateGT(y, a, polySieveM(X), s,
        treatProb = treatprob, nuisMod = "rlm",
        estimator = "ACW", target = "generalize", noi = F
)
acwfit %>% summary

```

We do better than the naive approach in all these cases.
