---
title: "Covariate Adjustment using ML predictions"
author: Apoorva Lal
date: "`r Sys.Date()`"
---

We simulate some experimental data with effect heterogeneity, and
use the Guo et al (2021) `mlrate` approach to estimate treatment effects using regression adjustment.

```{r}
suppressPackageStartupMessages(library(causalTransportR))
set.seed(42)

# simulate RCT
treatprob <- 0.5

# workhorse
dgp <- function(n = 1000, p = 10, treat.prob = treatprob,
                # bounds of X
                Xbounds = c(-1, 1),
                # nonlinear heterogeneity
                tauF = function(x) 1 / exp(-x[3]),
                # nonlinear y0
                y0F = function(x) x[1] + x[2] + sin(x[5]) + pmax(x[7], 0.5)) {
    X <- matrix(runif(n * p, Xbounds[1], Xbounds[2]), n, p)
    a <- rbinom(n, 1, treat.prob)
    # generate outcomes using supplied functions
    TAU <- apply(X, 1, tauF)
    Y0 <- apply(X, 1, y0F)
    # outcome
    y <- (a * TAU + Y0 + rnorm(n))
    cat("true effect \n")
    cat(mean(TAU), "\n")
    list(y = y, a = a, X = X)
}
```

## naive difference in means

```{r}
library(estimatr) # for linear regression with robust SEs
effect_se <- function(x) {
    x %>%
        summary() %>%
        .$coefficients %>%
        .[2, 1:2]
}

df <- dgp()
lm_robust(df$y ~ df$a) %>% effect_se()
```

## linear covariate adjustment

```{r}
lm_robust(df$y ~ df$a + df$X) %>% effect_se()
```

## ML covariate adjustment

Adjust for linear and interaction terms with LASSO.

```{r}
mlRate(df$y, df$a, polySieveM(df$X, k = 2, m = 2), nuisMod = "rlm")
```


__References__

Guo, Y., D. Coey, M. Konutgan, W. Li, C. Schoener, and M. Goldman. (2021): “Machine Learning for Variance Reduction in Online Experiments,” arXiv [stat.ML]
